### 2019-10-12 ch1_introduction/content.tex
new modifications

% CVC: weggewerkt --> KDP: niet alleen TFs maar volledige epigenetica regelt of een gen al dan niet tot expressie wordt gebracht, dus misschien kort vermelden dat het toch iets complexer is

@@

% CVC --> comment verwerkt. KDP: antigen-antibody: eerder over protein spreken ipv over antigen dan wordt het toch iets duidelijker, denk ik: een antibody herkent en bindt een bepaald proteine/eiwit,...
IHC (and many other biotechnologies) visualises antigen-antibody reactions by attaching particular molecules to the antibody, such as an enzyme that catalyses a colour-producing reaction, or a fluorescent chemical compound that can re-emit light upon excitation. The use of several colours (wavelengths) allows measuring expression levels of different antibodies simultaneously. Characterising cells in a semi-quantifiable way is labour intensive, however; since it involves acquiring an image of many cells and drawing a contour around each cell (called cell segmentation).
Modern implementations of IHC improve the throughput drastically by using robots and computer software to automate the image acquisition and cell segmentation. Nonetheless, the procedure is still labour intensive as the robots and computer software still needs to be kept in check.

Flow cytometry\cite{fulwyler_electronicseparationbiological_1965} is a technique which circumvents imaging and segmentation issues by having a steady stream of cells run through a laser and measuring the amount of light scattered from those cells. A major limitation of cytometry is that it only enables to study a suspension of viable single cells. Today, there are several protocols available to gain viable single cells from human tissues and tumours.

% Eventueel dit als ref: Leelatian N, Doxie DB, Greenplate AR, Sinnaeve J, Ihrie RA, Irish JM. Preparing viable single cells from human tissue and tumors for cytomic analysis. Current protocols in molecular biology. 2017 Apr;118(1):25C-1.
% --> CVC: Comment verwerkt KDP: circumvents segmentation: if you want to investigate tissues you have to dissociate the tissue . 
Flow cytometry technology enables to measure protein expression levels for millions of cells and tens of different antibodies. 
% KDP: vermeld hier hoeveel merkers flow cytometrie tegenwoordig aankan, want je komt er later op terug --> CVC zie mail

@@

\section{Computational tools}
Whole-genome profiling at single-cell level allows new types of analyses with which to study cellular heterogeneity at a hitherto unseen throughput. 
The new types of analyses permitted by single-cell omics present several computational challenges\cite{stegle_computationalanalyticalchallenges_2015, yuan_challengesemergingdirections_2017, chen_singlecellrnaseqtechnologies_2019}. This necessitates the development of novel computational tools, either because the problem statement of the performed analysis is completely novel, or to adapt existing methodology to new data characteristics (e.g. dimensionality and noise levels). Over the past five years, already 450 new tools for analysing single-cell omics data have been developed\cite{zappia_exploringsinglecellrnaseq_2018}. These tools typically fall within one of nine categories (Figure \ref{fig:comp_tools}). We discuss each category in the following subsections.

@@

	\centering\fontsize{9}{11}\selectfont
	\begin{tabularx}{\linewidth}{|p{2cm}llp{1.5cm}X|}
		\hline
		Name & Role & Host & Downloads per year & Description \\ \hline\hline
		\cranpkg{babelwhale} & aut & CRAN & 3996 & Interacting with Docker and Singularity containers \\
		\cranpkg{diffusionMap} & aut & CRAN & 21'361 & Implements diffusion map method of data parameterization, including creation and visualization of diffusion map \\
		\githubpkg{dynverse}{dynbenchmark} & aut & Github & \notavailable & Pipeline for benchmarking trajectory inference methods \\
		\cranpkg{dyndimred} & aut & CRAN & 5511 & Applying dimensionality reduction methods \\
		\githubpkg{dynverse}{dyneval} & aut & Github & \notavailable & Evaluating trajectory inference methods \\
		\githubpkg{dynverse}{dynfeature} & aut & Github & \notavailable & Calculating feature importance scores from trajectories \\
		\githubpkg{dynverse}{dyngen} & aut & Github & \notavailable & Simulating single-cell data using gene regulatory networks \\
		\githubpkg{dynverse}{dynguidelines} & aut & Github & \notavailable & User guidelines for trajectory inference \\
		\githubpkg{dynverse}{dynmethods} & aut & Github & \notavailable & A collection of wrappers for trajectory inference methods \\
		\githubpkg{dynverse}{dyno} & aut & Github & \notavailable & A pipeline for inferring, visualising and interpreting trajectories \\
		\cranpkg{dynparam} & aut & CRAN & 3084 & Creating meta-information for parameters \\
		\githubpkg{dynverse}{dynplot} & aut & Github & \notavailable & A simple visualisation library for trajectories \\
		\githubpkg{dynverse}{dynplot2} & aut & Github & \notavailable & A fully customisable visualisation library for trajectories \\
		\githubpkg{dynverse}{dyntoy} & aut & Github & \notavailable & Generating simple toy data of cellular differentiation \\
		\cranpkg{dynutils} & aut & CRAN & 5657 & Common functionality for the dynverse packages \\
		\githubpkg{dynverse}{dynwrap} & aut & Github & \notavailable & A common format for trajectories \\
		\cranpkg{GillespieSSA} & aut & CRAN & 7546 & Gillespie's Stochastic Simulation Algorithm (SSA) \\
		\cranpkg{GillespieSSA2} & aut & CRAN & 6506 & Gillespie's Stochastic Simulation Algorithm for Impatient People \\
		\githubpkg{dynverse}{gng} & aut & Github & \notavailable & An Rcpp implementation of the Growing Neural Gas algorithm \\
		\cranpkg{incgraph} & aut & CRAN & 3175 & Incremental graphlet counting for network optimisation \\
		\cranpkg{lmds} & aut & CRAN & \notavailable & Landmark Multi-Dimensional Scaling \\
		\cranpkg{princurve} & aut & CRAN & 26'991 & Fits a principal curve in arbitrary dimension \\
		\cranpkg{proxyC} & aut & CRAN & 117'484 & Computes proximity in large sparse matrices \\
		\cranpkg{qsub} & aut & CRAN & 3193 & Running commands remotely on gridengine clusters \\
		\cranpkg{SCORPIUS} & aut & CRAN & 4772 & Inferring developmental chronologies from single-cell RNA sequencing data \\ \hline\hline

>>>>

@@

IHC (and many other biotechnologies) visualises antigen-antibody reactions by attaching particular molecules to the antibody, such as an enzyme that catalyses a colour-producing reaction, or a fluorescent chemical compound that can re-emit light upon excitation. The use of several colours (wavelengths) allows measuring expression levels of different antibodies simultaneously. Characterising cells in a semi-quantifiable way is labour intensive, however; since it involves acquiring an image of many cells and drawing a contour around each cell (called cell segmentation).
Modern implementations of IHC improve the throughput drastically by using robots and computer software to automate the image acquisition and cell segmentation. Nonetheless, the procedure is still labour intensive as the robots and computer software still needs to be kept in check.

Flow cytometry\cite{fulwyler_electronicseparationbiological_1965} is a technique which circumvents imaging and segmentation issues by having a steady stream of cells run through a laser and measuring the amount of light scattered from those cells. 
Since cells need to be suspended in a buffer, flow cytometry is particularly useful for analysing non-adherent cells such as the many different immune cells in the blood. However, many protocols already exist to extract viable single cells from tissue as well\cite{leelatian_preparingviablesingle_2017}.
Common flow cytometry enables to measure proteins expression levels of millions of cells for 8 different antibodies, though state-of-the-art instrumentation allow detecting up to 27 biomarkers simultaneously\cite{cossarizza_guidelinesuseflow_2017}.

@@

\section{Computational tools}
Whole-genome profiling at single-cell level allows new types of analyses with which to study cellular heterogeneity at a hitherto unseen throughput. 
The new types of analyses permitted by single-cell omics present several computational challenges\cite{stegle_computationalanalyticalchallenges_2015, yuan_challengesemergingdirections_2017, chen_singlecellrnaseqtechnologies_2019}. This necessitates the development of novel computational tools, either because the problem statement of the performed analysis is completely novel, or to adapt existing methodology to new data characteristics (e.g. dimensionality and noise levels). Over the past five years, already 450 new tools for analysing single-cell omics data have been developed\cite{zappia_exploringsinglecellrnaseq_2018}. These tools typically fall within one of nine categories (Figure \ref{fig:comp_tools}). We discuss each category in the following subsections.
Whole-genome profiling at a single-cell level harbours exceptional opportunities to discover new aspects of biology and redefine existing knowledge.
% KDP: the new type of computational analysis etc... of eerder zeggen dat met de ontwikkeling van single cell technieken ook nieuwe computationele methoden moe(s)ten ontwikkeld worden
(Figure \ref{fig:comp_tools})

@@

	\centering\fontsize{9}{11}\selectfont
	\begin{tabularx}{\linewidth}{|p{2cm}llp{1.5cm}X|}
		\hline
		Name & Role & Host & Downloads per year & Description \\ \hline\hline
		\cranpkg{babelwhale} & aut & CRAN & 3996 & Interacting with Docker and Singularity containers \\
		\cranpkg{diffusionMap} & aut & CRAN & 21'361 & Implements diffusion map method of data parameterization, including creation and visualization of diffusion map \\
		\githubpkg{dynverse}{dynbenchmark} & aut & Github & \notavailable & Pipeline for benchmarking trajectory inference methods \\
		\cranpkg{dyndimred} & aut & CRAN & 5511 & Applying dimensionality reduction methods \\
		\githubpkg{dynverse}{dyneval} & aut & Github & \notavailable & Evaluating trajectory inference methods \\
		\githubpkg{dynverse}{dynfeature} & aut & Github & \notavailable & Calculating feature importance scores from trajectories \\
		\githubpkg{dynverse}{dyngen} & aut & Github & \notavailable & Simulating single-cell data using gene regulatory networks \\
		\githubpkg{dynverse}{dynguidelines} & aut & Github & \notavailable & User guidelines for trajectory inference \\
		\githubpkg{dynverse}{dynmethods} & aut & Github & \notavailable & A collection of wrappers for trajectory inference methods \\
		\githubpkg{dynverse}{dyno} & aut & Github & \notavailable & A pipeline for inferring, visualising and interpreting trajectories \\
		\cranpkg{dynparam} & aut & CRAN & 3084 & Creating meta-information for parameters \\
		\githubpkg{dynverse}{dynplot} & aut & Github & \notavailable & A simple visualisation library for trajectories \\
		\githubpkg{dynverse}{dynplot2} & aut & Github & \notavailable & A fully customisable visualisation library for trajectories \\
		\githubpkg{dynverse}{dyntoy} & aut & Github & \notavailable & Generating simple toy data of cellular differentiation \\
		\cranpkg{dynutils} & aut & CRAN & 5657 & Common functionality for the dynverse packages \\
		\cranpkg{dynwrap} & aut & CRAN & ? & A common format for trajectories \\
		\cranpkg{GillespieSSA} & aut & CRAN & 7546 & Gillespie's Stochastic Simulation Algorithm (SSA) \\
		\cranpkg{GillespieSSA2} & aut & CRAN & 6506 & Gillespie's Stochastic Simulation Algorithm for Impatient People \\
		\githubpkg{dynverse}{gng} & aut & Github & \notavailable & An Rcpp implementation of the Growing Neural Gas algorithm \\
		\cranpkg{incgraph} & aut & CRAN & 3175 & Incremental graphlet counting for network optimisation \\
		\cranpkg{lmds} & aut & CRAN & \notavailable & Landmark Multi-Dimensional Scaling \\
		\cranpkg{princurve} & aut & CRAN & 26'991 & Fits a principal curve in arbitrary dimension \\
		\cranpkg{proxyC} & aut & CRAN & 117'484 & Computes proximity in large sparse matrices \\
		\cranpkg{qsub} & aut & CRAN & 3193 & Running commands remotely on gridengine clusters \\
		\cranpkg{SCORPIUS} & aut & CRAN & 4772 & Inferring developmental chronologies from single-cell RNA sequencing data \\ \hline\hline
